### 2015-06-18 chapters/ipa_meets_unicode.tex
update different notions of diacritics

   \item The uses of the apostrophe has led to long discussions on the Unicode
        Standard email list. An English keyboard inputs <{\fontspec{Monaco}'}>
        \textsc{apostrophe} at \uni{0027}, although the preferred Unicode
        apostrophe is the <\ ' > \textsc{right single quotation mark} at
        \uni{2019}.\footnote{Note that many word processors (like Microsoft
        Word) by default will replace straight quotes by curly quotes,
        depending on the whitespace around it.} However, the glottal
        stop/glottalization/ejective marker is yet another completely different
        character <{\large ʼ}>, the \textsc{modifier letter apostrophe} at
        \uni{02BC}, which unfortunately looks mostly extremely similar to
        \uni{2019}. 
  \item Another problem is the <ˁ> \textsc{modifier letter reversed glottal
        stop} at \uni{02C1} vs.\@ the <ˤ> \textsc{modifier letter small reversed
        glottal stop} at \uni{02E4}. Both appear in various resources
        representing phonetic data online. This is thus a clear example for
        which the Unicode Standard does not solve the linguistic standardization
        problem.
  \item Even linguists are unlikely to distinguish between the <ə>
        \textsc{latin small letter schwa} at code point \uni{0259} and <ǝ>
        \textsc{latin small letter turned e} at \uni{01DD}.
  \item The alveolar click <ǃ> at \uni{01C3} is of course often simply
        typed as <!> \textsc{exclamation mark} at \uni{0021}.\footnote{In the
        Unicode Standard the <ǃ> at \unif{01C3} is labeled \textsc{latin letter
        retroflex click}, but in IPA that glyph is used for an alveolar or
        postalveolar click (not retroflex). This naming is probably best seen as
        an error in the Unicode Standard.}
  \item The dental click <ǀ>, in Unicode known as \textsc{latin letter dental
        click} at \uni{01C0}, is of course often simply typed as <|>
        \textsc{vertical line} at \uni{007C}.
  \item For the marking of length there is a special Unicode characters, namely
        <ː> \textsc{modifier letter triangular colon} at \uni{02D0}. However,
        typing <:> \textsc{colon} at \uni{003A} is of course much easier.

@@

@@

% problem with tie bar!!!

In Section~\ref{pitfall-different-notions-of-diacritics}, we highlight
the division of linguistic diacritics into two categories in the Unicode
Standard: Spacing Modifier Letters and Combining Diacritical Marks.\footnote{The
former contain characters like aspiration < ʰ > which is user specified (before
or after the base character) and the latter which combines with the base
character.} The tie-bar is somewhere in between these two character classes,
i.e.~when the combining breve (above or below) is told to combine within a
sequence of Unicode characters, it will bind to the character immediately
preceding it -- the so-called base character -- regardless if that base
character is a Spacing Modifier Letter \textit{or} Combining Diacritical Mark.
\footnote{In some cases conversion of a string sequence into Unicode
Normalization Form C alleviates the problem by composing sequences of characters
like <e> + < ́ > into single character < é > (see however diacritics like the
rhotic hook in ).}

>>>>

   \item The uses of the apostrophe has led to long discussions on the Unicode
        Standard email list. An English keyboard inputs <{\fontspec{Monaco}'}>
        \textsc{apostrophe} at \uni{0027}, although the preferred Unicode
        apostrophe is the <\ ' > \textsc{right single quotation mark} at
        \uni{2019}.\footnote{Note that many word processors (like Microsoft
        Word) by default will replace straight quotes by curly quotes,
        depending on the whitespace around it.} However, the glottal
        stop/glottalization/ejective marker is yet another completely different
        character <{\large ʼ}>, namely the \textsc{modifier letter apostrophe} at
        \uni{02BC}, which unfortunately looks mostly extremely similar to
        \uni{2019}. 
  \item Another problem is the <ˁ> \textsc{modifier letter reversed glottal
        stop} at \uni{02C1} vs.\@ the <ˤ> \textsc{modifier letter small reversed
        glottal stop} at \uni{02E4}. Both appear in various resources
        representing phonetic data online. This is thus a clear example for
        which the Unicode Standard does not solve the linguistic standardization
        problem.
  \item Even linguists are unlikely to distinguish between the <ə>
        \textsc{latin small letter schwa} at code point \uni{0259} and <ǝ>
        \textsc{latin small letter turned e} at \uni{01DD}.
  \item The alveolar click <ǃ> at \uni{01C3} is of course often simply
        typed as <!> \textsc{exclamation mark} at \uni{0021}.\footnote{In the
        Unicode Standard the <ǃ> at \unif{01C3} is labeled \textsc{latin letter
        retroflex click}, but in IPA that glyph is used for an alveolar or
        postalveolar click (not retroflex). This naming is probably best seen as
        an error in the Unicode Standard.}
  \item The dental click <ǀ>, in Unicode known as \textsc{latin letter dental
        click} at \uni{01C0}, is of course often simply typed as <|>
        \textsc{vertical line} at \uni{007C}.
  \item For the marking of length there is a special Unicode characters, namely
        <ː> \textsc{modifier letter triangular colon} at \uni{02D0}. However,
        typing <:> \textsc{colon} at \uni{003A} is of course much easier.

@@

% this has to be revised: there are two issues: homoglyphs and alternative 
encoding options. that is not necessarily the same. Also include minor group break 
vs. dental click problem. and stress vs. ejective?

@@

Further, the IPA tie bar (see Section~\ref{pitfall-tie-bar}) is somewhere in
between these two character classes. The Unicode encoding of this character is
uses a Combining Diacritical Mark, namely either \textsc{combining double
inverted breve} at \uni{0361} or \textsc{combining double breve below} at
\uni{035C}. Such a combining mark is by definition tied to the character in
front, but not the character following it. The Unicode treatment of this
character thus only partly corresponds to the IPA conception, which ideally
would have the tie bar linked both to the character in front and to the 
character following.
